---
title: "lab_6"
author: "Drew Andersen"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```


## Rank based test example

Mann-Whitney U (non-parametric alternative to an unpiared 1-sample t-test)
H0: Equal ranks(or medians) - "perfectly intermingeled"
HA: Ranks are not equal (or, medians are not equal)

### Create samples
`r set.seed()` to create "pseudorandom, but not reproducible sample"

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at distributions with hist():
```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results:
```{r}
ttest <- t.test(gp_1, gp_2)
#If these data were paired, I can just add the argument (paired = "TRUE"), within the t-test
```

This p-value means that IF our samples are drawn from populations with the same mean (if null hyp is true), there is a `r round(ttest$p.value,2)*100`% probability of finding samples mean at least as different as ours by random chance.

[A more comprhensive explannation of difference between the groups...actually means, data spread, effect size]... there is no sig diff (t(df)=t-stat value, p = ).

### Mann Whitney U rank based test

We use the `wilcox.test()` function to perform a two-sample rank-based test. 
```{r}
mwu <- wilcox.test(gp_1, gp_2)
```
use the same def as above, but replace "mean" with "median"
- no sig siff in medians (Mann-Whitney U; U = `r mwu$statistic`, p = `r round(mwu$p.value, 2)`)

